An Extendable \proglang{Python} Implementation of Robust Optimisation Monte Carlo

### 2022-02-02 latex_files/ch2/2-1-implicit-models.tex
reordered files and fixed graph colors
Add comments and some changes. Ask HP.

An implicit or simulator-based model is a parameterised stochastic
data generating mechanism. The key characteristic of these models is
that we can sample, or simulate, data points, but that we cannot
evaluate the likelihood. Formally, a simulator-based model is a
parameterised family of probability density functions $\{
p(\yb|\thetab) \}_{\thetab}$, whose closed-form is either unknown or
computationally intractable. Whereas evaluating $p(\yb|\thetab)$ is
intractable, sampling is feasible. Practically, a simulator can be
understood as a random black-box machine (computer code) that
generates samples $\yb$ in a stochastic manner for any given a set of
parameters $\thetab$. We can isolate the randomness by introducing the
stochastic nuisance variables $\vb \sim p(\vb)$ so that the simulator
becomes a deterministic mapping $\simulator$ that maps
$(\thetab, \vb)$ to the data $\yb$, i.e.\
$\yb=\simulator(\thetab,\vb)$. The distribution $p(\vb)$ is determined
by the random draws performed when running the computer
code. Importantly, we can fix the value of $\vb$ while changing
$\thetab$ by fixing the random seed used.

Simulator-based models provide considerable modelling freedom; any
physical process that can be conceptualised as a computer program of
finite steps can be modelled as a simulator-based model without any
compromise. The modelling freedom allows for any amount of hidden
(unobserved) internal variables or rule-based decisions. Hence,
implicit models are often used to model physical phenomena in the
natural sciences such as genetics, epidemiology, or neuroscience, to
name a few examples. Further background on simulator-based models and
example applications can be found in the articles
by \citet{Gutmann2016, Lintusaari2017, Sisson2018, Cranmer2020}.

The modelling freedom of simulator-based models, however, comes at the
price of difficulties in inferring their parameters. Denoting observed
data by $\data$, the main difficulty is that the likelihood function
$L(\thetab) = p(\data|\thetab)$ is intractable. To better see the
sources of the intractability, and to address them, we go back to the
basic characterisation of the likelihood as the (rescaled) probability
for a parameter of the model to generate data $\yb$ that is similar to
the observed data $\data$. More formally, the likelihood $L(\thetab)$
equals
\begin{equation} \label{eq:likelihood}
  L(\thetab) = \lim_{\epsilon \to 0} c_\epsilon \int_{\yb \in B_{d,\epsilon}(\data)} p(\yb|\thetab)d\yb =
  \lim_{\epsilon \to 0} c_\epsilon \Pr(\simulator(\thetab, \vb) \in \region(\data))
\end{equation}
where $c_\epsilon$ is a proportionality factor that depends on
$\epsilon$ and $\region(\data)$ is an $\epsilon$ region around $\data$
that is defined via a distance function $d$, i.e.\ $\region(\data)
:= \{\yb: d(\yb, \data) \leq \epsilon \}$.

The basic characterisation of the likelihood in \eqref{eq:likelihood}
highlights two sources of intractability: The first is the computation
of the probability $\Pr(\simulator(\thetab,\vb) \in \region(\data))$, the second
is the limit of $\epsilon \to 0$.  Whilst the probability can be
approximated with samples, this becomes computationally infeasible if
$\epsilon$ is too small. Hence, a large class of inference methods
work with $\epsilon >0$, which leads to the approximate likelihood
function $L_{d, \epsilon}(\thetab)$,
\begin{equation} \label{eq:approx_likelihood}
  L_{d, \epsilon}(\thetab) = \Pr(\yb \in \region(\data)), \quad \text{where  } \epsilon > 0.
\end{equation}
Most often in likelihood-free inference, we are interested in the posterior distribution of the parameters $\thetab$ given some observed data. The approximate likelihood function in \eqref{eq:approx_likelihood} implies the approximate posterior  $p_{d,\epsilon}(\thetab|\data)$, 
\begin{equation} \label{eq:approx_posterior}
  p_{d,\epsilon}(\thetab|\data) \propto L_{d, \epsilon}(\thetab) p(\thetab).
\end{equation}

>>>>

An implicit or simulator-based model is a parameterised stochastic
data generating mechanism. The key characteristic of these models is
that we can sample, or simulate, data points, but that we cannot
evaluate the likelihood. Formally, a simulator-based model is a
parameterised family of probability density functions $\{
p(\yb|\thetab) \}_{\thetab}$, whose closed-form is either unknown or
computationally intractable. Whereas evaluating $p(\yb|\thetab)$ is
intractable, sampling is feasible. Practically, a simulator can be
understood as a random black-box machine (computer code) that
generates samples $\yb$ in a stochastic manner for any given a set of
parameters $\thetab$. We can isolate the randomness by introducing the
stochastic nuisance variables $\vb \sim p(\vb)$ so that the simulator
becomes a deterministic mapping $\simulator$ that maps
$(\thetab, \vb)$ to the data $\yb$, i.e.\
$\yb=\simulator(\thetab,\vb)$. The distribution $p(\vb)$ is determined
by the random draws performed when running the computer
code. \textbf{HP comment: I don't think that the
distribution $p(\vb)$ is determined by the random draws, but $p(\vb)$ 
should define the way we choose the `random' draws. How about: The distribution 
$p(\vb)$ is defined as a random number generating process within the computer 
code.} Importantly, we can fix the value of $\vb$ while changing
$\thetab$ by fixing the random seed used.

Simulator-based models provide considerable modelling freedom; any
physical process that can be conceptualised as a computer program of
finite steps can be modelled as a simulator-based model without any
compromise. The modelling freedom allows for any amount of hidden
(unobserved) internal variables or rule-based decisions. Hence,
implicit models are often used to model physical phenomena in the
natural sciences such as e.g.~genetics, epidemiology, or neuroscience. 
Further background on simulator-based models and
example applications can be found in the articles
by \citet{Gutmann2016, Lintusaari2017, Sisson2018, Cranmer2020}.

The modelling freedom of simulator-based models, however, comes at the
price of difficulties in inferring their parameters. Denoting observed
data by $\data$, the main difficulty is that the likelihood function
$L(\thetab) = p(\data|\thetab)$ is generally intractable. To better see the
sources of the intractability, and to address them, we go back to the
basic characterisation of the likelihood as the (rescaled) probability
of a parameter of the model to generate data $\yb$ that is similar to
the observed data $\data$. More formally, the likelihood $L(\thetab)$
equals
\begin{equation} \label{eq:likelihood}
  L(\thetab) = \lim_{\epsilon \to 0} c_\epsilon \int_{\yb \in B_{d,\epsilon}(\data)} p(\yb|\thetab)d\yb =
  \lim_{\epsilon \to 0} c_\epsilon \Pr(\simulator(\thetab, \vb) \in \region(\data)  \mid \theta)
\end{equation}
where $c_\epsilon$ is a proportionality factor that depends on
$\epsilon$ and $\region(\data)$ is an $\epsilon$ region around $\data$
that is defined via a distance function $d$, i.e.\ $\region(\data)
:= \{\yb: d(\yb, \data) \leq \epsilon \}$.

The basic characterisation of the likelihood in \eqref{eq:likelihood}
highlights two sources of intractability: The first is the computation
of the probability $\Pr(\simulator(\thetab,\vb) \in \region(\data))$,
the second is the limit of $\epsilon \to 0$.  Whilst the probability
can be approximated with samples, this becomes computationally
infeasible if $\epsilon$ is too small. Hence, a large class of
inference methods work with $\epsilon >0$, which leads to the
approximate likelihood function $L_{d, \epsilon}(\thetab)$,
\begin{equation} \label{eq:approx_likelihood}
  L_{d, \epsilon}(\thetab) = \Pr(\yb \in \region(\data) \mid \theta), \quad \text{where  } \epsilon > 0.
\end{equation}
Most often in likelihood-free inference, we are interested in the
posterior distribution of the parameters $\thetab$ given some observed
data. \textbf{HP comment: thinking about it, I think we're often more 
interested in a sample from the posterior, but this is maybe just semantics.} The approximate likelihood function in
\eqref{eq:approx_likelihood} implies the approximate posterior
$p_{d,\epsilon}(\thetab|\data)$,
\begin{equation} \label{eq:approx_posterior}
  p_{d,\epsilon}(\thetab|\data) \propto L_{d, \epsilon}(\thetab) p(\thetab).
\end{equation}

### 2022-02-02 latex_files/ch2/2-2-OMC.tex
reordered files and fixed graph colors
Add comments and some changes. Ask HP.

Our description of OMC \citep{Meeds2015}
follows \citet{Ikonomov2019} who base their explanation of OMC on the
approximate likelihood function in \eqref{eq:approx_likelihood}. With
the indicator function
%
\begin{equation} \label{eq:indicator} \indicator{\region(\yb)}(\xb)=
  \left\{
    \begin{array}{ll}
      1 & \mbox{if } \xb \in \region(\yb) \\
      0 & \mbox{otherwise} 
    \end{array} \right. \end{equation}
%
we can write the approximate likelihood function $L_{d, \epsilon}(\thetab)$ in \eqref{eq:approx_likelihood} as
\begin{gather} \label{eq:approx_likelihood_omc}
  L_{d, \epsilon}(\thetab) = \Pr(\yb \in \region(\data)) =
  \int_{\yb \in B_{d,\epsilon}(\data)}p(\yb|\thetab)d\yb =
  \int_{\yb \in \R^D} \indicator{\region(\data)}(\yb)p(\yb|\thetab)d\yb
\end{gather}
which we can approximate as a sample average
\begin{equation}
L_{d, \epsilon}(\thetab) \approx \frac{1}{N} \sum_{i=1}^N \indicator{\region (\data)} (\yb_i)
 \quad \text{ where } \yb_i = \simulator(\thetab, \vb_i),\; \vb_i \sim p(\vb). \label{eq:alt_view}
\end{equation}
We thus approximate the likelihood of a specific $\thetab$ by counting
the fraction of samples that lie inside the area around the
observations. The fact that we isolated the randomness of the
simulator via the $\vb_i$ has two crucial consequences: First, we can
sample the nuisance variables $\vb_i \sim p(\vb)$ only once and use
the same set of samples to approximate $L_{d,\epsilon}(\thetab)$ for
different values of $\thetab$. Secondly, since
$\yb_i=\simulator(\thetab, \vb_i)$ is a function of $\thetab$,
checking whether $\yb_i$ is contained in region $\region (\data)$ for
a particular value of $\thetab$ is the same as checking whether
$\thetab$ is in the acceptance region $\accregioni
= \{ \thetab: \simulator(\thetab, \vb_i) \in \region(\data) \}$. This
leads to the likelihood approximation
\begin{equation}
L_{d, \epsilon}(\thetab) \approx \frac{1}{N} \sum_{i=1}^N \indicator{\accregioni}(\thetab) \label{eq:alt_view_theta}
\end{equation}
and the corresponding approximate posterior
\begin{equation} \label{eq:approx_posterior_omc}
  p_{d,\epsilon}(\thetab|\data) \propto
  p(\thetab) \sum_i^N  \indicator{\accregioni}(\thetab).
\end{equation}
As argued by \citet{Ikonomov2019}, these derivations provide a unique
perspective for likelihood-free inference by shifting the focus onto
the geometry of the acceptance regions $\accregioni$. Indeed, the task
of approximating the likelihood and/or posterior becomes a task of
characterising the sets $\accregioni$.

OMC by \citet{Meeds2015} assumes that the distance $d$ is the
Euclidean distance between summary statistics $\Phi$ of the observed
and generated data, and that the $\accregioni$ can be well
approximated by infinitesimally small ellipses. These assumptions leads to an approximation of the posterior in
terms of weighted samples $\thetab_i^*$ that achieve the smallest
distance between observed and simulated data for each realisation
$\vb_i \sim p(\vb)$, i.e.\
\begin{equation} \label{eq:omc_opt_prob}
\thetab_i^* = \argmin_{\thetab} ||\Phi(\data)-\Phi(\simulator(\thetab, \vb_i))||_2  , \quad \vb_i \sim p(\vb).
\end{equation}
The weighting for each $\thetab_i^*$ is proportional to the prior
density at $\thetab_i^*$ and inversely proportional to the determinant
of the Jacobian matrix of the summary statistics at $\thetab_i^*$. For
further details on OMC we refer the reader to \citep{Meeds2015,
Ikonomov2019}.

>>>>

Our description of OMC \citep{Meeds2015} follows \citet{Ikonomov2019}
who base their explanation of OMC on the approximate likelihood
function in \eqref{eq:approx_likelihood}. With the indicator function
%
\begin{equation} \label{eq:indicator} \indicator{\region(\yb)}(\xb)=
  \left\{
    \begin{array}{ll}
      1 & \mbox{if } \xb \in \region(\yb) \\
      0 & \mbox{otherwise} 
    \end{array} \right. \end{equation}
%
we can write the approximate likelihood function $L_{d, \epsilon}(\thetab)$ in \eqref{eq:approx_likelihood} as
\begin{gather} \label{eq:approx_likelihood_omc}
  L_{d, \epsilon}(\thetab) = \Pr(\yb \in \region(\data)) =
  \int_{\yb \in B_{d,\epsilon}(\data)}p(\yb|\thetab)d\yb =
  \int_{\yb \in \R^D} \indicator{\region(\data)}(\yb)p(\yb|\thetab)d\yb
\end{gather}
which we can approximate as a sample average
\begin{equation}
L_{d, \epsilon}(\thetab) \approx \frac{1}{N} \sum_{i=1}^N \indicator{\region (\data)} (\yb_i)
 \quad \text{ where } \yb_i = \simulator(\thetab, \vb_i),\; \vb_i \sim p(\vb). \label{eq:alt_view}
\end{equation}
We thus approximate the likelihood of a specific $\thetab$ by counting
the fraction of samples that lie inside the volume around the
observations. The fact that we isolated the randomness of the
simulator via the $\vb_i$ has two crucial consequences: First, we can
sample the nuisance variables $\vb_i \sim p(\vb)$ only once and use
the same set of samples to approximate $L_{d,\epsilon}(\thetab)$ for
different values of $\thetab$. Secondly, since
$\yb_i=\simulator(\thetab, \vb_i)$ is a function of $\thetab$,
checking whether $\yb_i$ is contained in region $\region (\data)$ for
a particular value of $\thetab$ is the same as checking whether
$\thetab$ is in the acceptance region $\accregioni
= \{ \thetab: \simulator(\thetab, \vb_i) \in \region(\data) \}$. This
leads to the likelihood approximation
\begin{equation}
L_{d, \epsilon}(\thetab) \approx \frac{1}{N} \sum_{i=1}^N \indicator{\accregioni}(\thetab) \label{eq:alt_view_theta}
\end{equation}
and the corresponding approximate posterior
\begin{equation} \label{eq:approx_posterior_omc}
  p_{d,\epsilon}(\thetab|\data) \propto
  p(\thetab) \sum_i^N  \indicator{\accregioni}(\thetab).
\end{equation}
As argued by \citet{Ikonomov2019}, these derivations provide a unique
perspective for likelihood-free inference by shifting the focus onto
the geometry of the acceptance regions $\accregioni$. Indeed, the task
of approximating the likelihood and/or posterior becomes a task of
characterising the sets $\accregioni$.

OMC by \citet{Meeds2015} assumes that the distance $d$ is the
Euclidean distance $||\cdot||_2$ between summary statistics $\Phi$ of the observed
and generated data, and that the $\accregioni$ can be well
approximated by infinitesimally small ellipses. These assumptions
leads to an approximation of the posterior in terms of weighted
samples $\thetab_i^*$ that achieve the smallest distance between
observed and simulated data for each realisation $\vb_i \sim p(\vb)$,
i.e.\
\begin{equation} \label{eq:omc_opt_prob}
\thetab_i^* = \argmin_{\thetab} ||\Phi(\data)-\Phi(\simulator(\thetab, \vb_i))||_2  , \quad \vb_i \sim p(\vb).
\end{equation}
The weighting for each $\thetab_i^*$ is proportional to the prior
density at $\thetab_i^*$ and inversely proportional to the determinant
of the Jacobian matrix of the summary statistics at $\thetab_i^*$. For
further details on OMC we refer the reader to \citep{Meeds2015,
  Ikonomov2019}.

### 2022-02-02 latex_files/ch2/2-3-ROMC.tex
Add comments and some changes. Ask HP.

\citet{Ikonomov2019} showed that considering infinitesimally small
ellipses can lead to highly overconfident posteriors. We refer the
reader to their paper for the technical details and conditions for
this issue to occur. Intuitively, it happens because the weights in
OMC are only computed from information at $\thetab_i^*$. And using
only local information can be misleading. For example if the curvature
of $||\Phi(\data)-\Phi(\simulator(\thetab, \vb_i))||_2$ at
$\thetab_i^*$ is nearly flat, the curvature alone may wrongly indicate
that $\accregioni$ is much larger than it actually is. We implement in
our software package the robust generalisation of OMC by
\citet{Ikonomov2019} that resolves this issue.

ROMC approximates the acceptance regions $\accregioni$ and defines
local proposal distributions $q_i(\thetab)$ on them from which the
posterior samples $\thetab_{ij} \sim q_i$ are generated. The samples
are assigned (importance) weights $w_{ij}$ that compensate for using
the proposal distributions $q_i(\thetab)$ and not the prior
$p(\thetab)$,
\begin{equation}
  w_{ij} = \frac{\indicator{\accregioni}(\thetab_{ij}) p(\thetab_{ij})}{q(\thetab_{ij})}.
  \label{eq:sampling}
\end{equation}
Given the weighted samples, any expectation
$\Ex_{p(\thetab|\data)}[h(\thetab)]$ of some function $h(\thetab)$, can be approximated as
\begin{equation} \label{eq:expectation}
  \Ex_{p(\thetab|\data)}[h(\thetab)] \approx \frac{\sum_{ij} w_{ij} h(\thetab_{ij})}{\sum_{ij} w_{ij}}
\end{equation}
\citet{Ikonomov2019} considered uniform distributions as proposal
distributions so that the main work is to approximate the acceptance
regions $\accregioni$ and to represent them such that uniform sampling
is easy. The approximation of the acceptance regions contains two
compulsory and one optional step: (1) solving the optimisation
problems as in OMC, (2) constructing bounding boxes around
$\accregioni$ and optionally, (3) refining the approximation via a
probabilistic surrogate model of the distance.
\subsubsection*{Solving the deterministic optimisation problems}
For each set of nuisance variables $\vb_i, i = \{1,2,\ldots,n_1 \}$, we
search for a point $\thetab^*_i$ such that
$d(\simulator(\thetab^*_i,\vb_i), \data) \le \epsilon$. For notational
convenience, we denote the distance $d(\simulator(\thetab,\vb_i), \data)$ by
$d_i(\thetab)$.  Obtaining $\theta_i^*$ involves solving the following
optimisation problem:
\begin{subequations}
\begin{alignat}{2}      
  &\!\min_{\thetab}        && d_i(\thetab) \label{eq:optProb}\\
  &\text{subject to} & \quad& d_i(\thetab) \leq \epsilon
\end{alignat}
\end{subequations}
%
The optimisation problem can be treated as unconstrained, accepting
the optimal point $\thetab_i^* = \text{argmin}_{\thetab} d_i(\thetab)$
only if $d_i(\thetab_i^*) \le \epsilon$. If $d_i(\thetab)$ is
differentiable any gradient-based optimizer can be used
for~\ref{eq:optProb}. The gradients $\nabla_{\thetab} d_i(\thetab)$
can be either provided in closed form or approximated by finite
differences. In case $d_i$ is not differentiable, Bayesian
Optimisation~\cite{Shahriari2016} provides an alternative choice. In
this scenario, apart from obtaining an optimal $\thetab_i^* $, a
surrogate model $\hat{d}_i(\thetab)$ of the distance function
$d_i(\thetab)$ is also automatically obtained; $\hat{d}_i$ can then
substitute the actual distance function in downstream steps of the
algorithms, which can lead to computational gains if running the
simulator and hence evaluating the distance function $d_i(\thetab)$ is
computationally expensive.

\subsubsection*{Estimating the acceptance regions}
The acceptance region $\accregioni$ is approximated by a bounding box
$\accregionihat$ around it. Ideally, we want the bounding box to be as
tight as possible to $\mathcal{S}_i$, to ensure high acceptance rate
in the importance sampling, and big enough for not discarding valid
parts of $\mathcal{S}_i$. The bounding boxes are built in two
steps. First, we define their axes $\mathbf{v}_d$ based on the
(estimated) curvature of the distance $\thetab_i^*$, second, we
determine the size of the box via a one-dimensional line-search method
along each axix, see Algorithm~\ref{alg:region_construction} for the
details. After the bounding boxes construction, a uniform distribution
$q_i$ is defined on each bounding box, and is used as the proposal
region for importance sampling.

@@

The training data for fitting the quadratic model is obtained by
sampling $\thetab \sim q_i$ and evaluating the corresponding distances
$d_i(\thetab)$. The generation of the training data adds an extra
computational cost, but leads to a significant speed-up when
evaluating the weights $w_{ij}$. Moreover, the extra cost is largely
eliminated if Bayesian Optimization with a Gaussian processe (GP)
surrogate model was used to obtain $\thetab_i^*$ in the first place
(see above). In this case, we can use the GP model of the distance
function instead of $d_i(\thetab)$ to generate the training data. This
essentially replaces the global GP model with a simpler local
quadratic model which is typically more robust.

>>>>

\citet{Ikonomov2019} showed that considering infinitesimally small
ellipses can lead to highly overconfident posteriors. We refer the
reader to their paper for the technical details and conditions for
this issue to occur. Intuitively, it happens because the weights in
OMC are only computed from information at $\thetab_i^*$, and using
only local information can be misleading. For example if the curvature
of $||\Phi(\data)-\Phi(\simulator(\thetab, \vb_i))||_2$ at
$\thetab_i^*$ is nearly flat, the curvature alone may wrongly indicate
that $\accregioni$ is much larger than it actually is. In
our software package we implement the robust generalisation of OMC by
\citet{Ikonomov2019} that resolves this issue.

ROMC approximates the acceptance regions $\accregioni$ and on them defines
local proposal distributions $q_i(\thetab)$ from which the
posterior samples $\thetab_{ij} \sim q_i$ are generated. The samples
are assigned (importance) weights $w_{ij}$ that compensate for using
the proposal distributions $q_i(\thetab)$ and not the prior
$p(\thetab)$,
\begin{equation}
  w_{ij} = \frac{\indicator{\accregioni}(\thetab_{ij}) p(\thetab_{ij})}{q(\thetab_{ij})}.
  \label{eq:sampling}
\end{equation}
Given the weighted samples, any expectation
$\Ex_{p(\thetab|\data)}[h(\thetab)]$ of some function $h(\thetab)$, can be approximated as
\begin{equation} \label{eq:expectation}
  \Ex_{p(\thetab|\data)}[h(\thetab)] \approx \frac{\sum_{ij} w_{ij} h(\thetab_{ij})}{\sum_{ij} w_{ij}}
\end{equation}
\citet{Ikonomov2019} considered uniform distributions as proposal
distributions so that the main task is to approximate the acceptance
regions $\accregioni$ and to represent them so that uniform sampling
is easy. The approximation of the acceptance regions contains two
compulsory and one optional step: (1) solving the optimisation
problems as in OMC, (2) constructing bounding boxes around
$\accregioni$ and optionally, (3) refining the approximation via a
probabilistic surrogate model of the distance.
\subsubsection*{Solving the deterministic optimisation problems}
For each set of nuisance variables $\vb_i, i = \{1,2,\ldots,n_1 \}$, we
search for a point $\thetab^*_i$ such that
$d(\simulator(\thetab^*_i,\vb_i), \data) \le \epsilon$. For notational
convenience, we denote the distance $d(\simulator(\thetab,\vb_i), \data)$ by
$d_i(\thetab)$.  Obtaining $\theta_i^*$ involves solving the following
optimisation problem:
\begin{subequations}
\begin{alignat}{2}      
  &\!\min_{\thetab}        && d_i(\thetab) \label{eq:optProb}\\
  &\text{subject to} & \quad& d_i(\thetab) \leq \epsilon
\end{alignat}
\end{subequations}
%
The optimisation problem can be treated as unconstrained, accepting
the optimal point $\thetab_i^* = \text{argmin}_{\thetab} d_i(\thetab)$
only if $d_i(\thetab_i^*) \le \epsilon$. If $d_i(\thetab)$ is
differentiable any gradient-based optimizer can be used
for~\ref{eq:optProb}. The gradients $\nabla_{\thetab} d_i(\thetab)$
can be either provided in closed form or approximated by finite
differences. In case $d_i$ is not differentiable, Bayesian
Optimisation~\citep{Shahriari2016} provides an alternative approach. In
this scenario, apart from obtaining an optimal $\thetab_i^* $, a
surrogate model $\hat{d}_i(\thetab)$ of the distance function
$d_i(\thetab)$ is also automatically obtained; $\hat{d}_i$ can then
substitute the actual distance function in downstream steps of the
algorithms, which can lead to computational gains if running the
simulator and hence evaluating the distance function $d_i(\thetab)$ is
computationally expensive.

\subsubsection*{Estimating the acceptance regions}
The acceptance region $\accregioni$ is approximated by a bounding box
$\accregionihat$ around it. Ideally, we want the bounding box to be as
tight as possible to $\mathcal{S}_i$, to ensure high acceptance rate
in the importance sampling, and big enough for not discarding valid
parts of $\mathcal{S}_i$. The bounding boxes are built in two
steps. First, we define their axes $\mathbf{v}_d$ based on the
(estimated) curvature of the distance $\thetab_i^*$. Second, we
determine the size of the box via a one-dimensional line-search method
along each axix, see Algorithm~\ref{alg:region_construction} for the
details. After the bounding boxes construction, a uniform distribution
$q_i$ is defined on each bounding box, and is used as the proposal
region for importance sampling.

@@

The training data for fitting the quadratic model is obtained by
sampling $\thetab \sim q_i$ and evaluating the corresponding distances
$d_i(\thetab)$. The generation of the training data adds an extra
computational cost, but leads to a significant speed-up when
evaluating the weights $w_{ij}$. Moreover, the extra cost is largely
eliminated if Bayesian Optimisation with a Gaussian process (GP)
surrogate model was used to obtain $\thetab_i^*$ in the first place
(see above \textbf{HP comment: what above this refers to?}). In this case, we can use the GP model of the distance
function instead of $d_i(\thetab)$ to generate the training data. This
essentially replaces the global GP model with a simpler local
quadratic model which is typically more robust.

### 2022-02-02 latex_files/ch2/2-4-ELFI.tex
reordered files and fixed graph colors
Add comments and some changes. Ask HP.

\pkg{Engine for Likelihood-Free Inference (ELFI)}\footnote{Extended
  documentation can be found
  \href{https://elfi.readthedocs.io/en/latest/}{here}}~\cite{1708.00707}
is a \proglang{Python} package for Likelihood-Free Inference
(LFI). We selected to implemented ROMC in \pkg{ELFI} since it provides
convenient modules for all the fundamental components of a
probabilistic model (e.g.\ prior, simulator, summaries
etc.). Furthermore, \pkg{ELFI} already supports some recently proposed
likelihood-free inference methods, making it straightforward to
perform comparisons between methods. The implementation was officially
added to \pkg{ELFI} in version \pkg{v0.8.0}.

@@

\pkg{ELFI} models the probabilistic model as a Directed Acyclic Graph
(DAG). This functionality is based on the package \pkg{NetworkX},
which supports general-purpose graphs. In most cases, the structure of
a likelihood-free model follows the pattern of figure~\ref{fig:elfi};
some edges connect the prior distributions to the simulator, the
simulator is connected to the summary statistics that, in turn, lead
to the output node. Samples can be obtained from all nodes through
sequential (ancestral) sampling. \pkg{ELFI} automatically considers as
parameters of interest, i.e. those we try to infer a posterior
distribution, as the ones defined through the \code{elfi.Prior}
class. Finally, the observations are passed to the appropriate node
through the argument
\code{observed}.

>>>>

\pkg{Engine for Likelihood-Free Inference (ELFI)}\footnote{Extended
  documentation can be found
  \href{https://elfi.readthedocs.io/en/latest/}{https://elfi.readthedocs.io}}~\cite{1708.00707}
is a \proglang{Python} package for LFI. We
selected to implemented ROMC in \pkg{ELFI} since it provides
convenient modules for all the fundamental components of a
probabilistic model (e.g.\ prior, simulator, summaries
etc.). Furthermore, \pkg{ELFI} already supports some recently proposed
likelihood-free inference methods, making it straightforward to
perform comparisons between methods. The implementation was officially
added to \pkg{ELFI} in version \pkg{v0.8.0}.

@@

\pkg{ELFI} handles the probabilistic model as a Directed Acyclic Graph
(DAG). This functionality is based on the package \pkg{NetworkX},
which supports general-purpose graphs. In most cases, the structure of
a likelihood-free model follows the pattern of Figure~\ref{fig:elfi};
some edges connect the prior distributions to the simulator, the
simulator is connected to the summary statistics that, in turn, lead
to the output node. Samples can be obtained from all nodes through
sequential (ancestral) sampling. \pkg{ELFI} automatically considers as
parameters of interest, i.e. those we try to infer a posterior
distribution, as the ones defined through the \code{elfi.Prior}
class. Finally, the observations are passed to the appropriate node
through the argument
\code{observed}.
